### 2016-04-30 icml_workshop/abstraction_desiderata.tex
Some commits

\section{Two Desiderata for Abstraction in RL}

We propose the following two desiderata on theories of abstraction in the context of planning or learning, given a Markov Decision Process, $M_G$:
\begin{enumerate}
\item \textsc{Compressibility}: To compress $M_G$ to be significantly smaller. \footnote{where smallness is understood as defined in the of \cite{LITTMAN COMPLEXITY}  blah blah blah}
\item \textsc{Optimality}: To induce an MDP $M_A$ such that solving $M_A$ basically solves $M_G$.
\end{enumerate}

\dnote{Defense of the desiderata}

Critically, either desiderata is trivial to satisfy on its own.

Given an \ac{MDP}, $M_G$, we can satisfy \textsc{Compressibility} by replacing $S_G$ with a state space that consists of a single (arbitrary) state from $S_G$. Clearly, such a resulting \ac{MDP} satisfies our first desiderata - we are left with an MDP that is substantially smaller than the original MDP.

Now consider \textsc{Optimality}. Again, we can develop a simple method for satisfying this desiderata on its own. Namely, consider the MDP $M_A$, that is identical to $M_G$. Thus, the optimal policy for $M_A$ is exactly the optimal policy for $M_G$, consequently satisfying \textsc{Optimality}.

The interesting case is satisfying {\it both} desiderata. To the best of our knowledge, no theory of temporal abstraction has been formally proven to satisfy the two - in the next section, we discuss prior methods of abstraction, and their relative status to these desiderata.