### 2016-04-30 icml_workshop/framework.tex
adding framy

\section{Abstraction Framework}

In previous work~\cite{abelhershko2016approx}, we extended the unifying framework of state abstraction proposed by~\cite{li2006towards}. In this work we develop a theory of approximate abstraction, in which abstraction is interpreted as state aggregation; that is, a state abstraction function $\phi : \mathcal{S}_G \mapsto \mathcal{S}_A$,

Formalism

Extension to Temp Ext

Satisfying Desiderata