### 2016-05-03 icml_workshop/abstract.tex

Abstraction is among the most fundamental tools to intelligent agents operating in complex spaces; through abstraction, intelligent agents may pare down otherwise prohibitively large problems into tractably solvable ones. Unsurprisingly, a great deal of work in planning and \ac{RL} has focused on abstraction in \ac{MDP} in the form of state aggregation and temporally extended actions. However, there exists little work detailing what good abstractions ought to formally accomplish. In this work we propose three desiderata on abstractions in the context of planning and \ac{RL} to motivate future research. We summarize existing work on unifying state aggregation methods, and discuss how our recent work in approximate state abstraction relates to these three desiderata.