### 2016-05-22 paper/paper.tex
prepping for merge

\documentclass{article}
\usepackage{times}
\usepackage{graphicx}
\usepackage{subfigure} 
\usepackage{natbib}
\usepackage{algorithm}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}%For stating numbered theorems
\usepackage{amsfonts}
\usepackage{algorithmic}
\usepackage{subfigure}
\usepackage{hyperref}
\newcommand{\theHalgorithm}{\arabic{algorithm}} % Fixes misbehavior of conflicting packages.
\usepackage[accepted] {icml2016/icml2016}
\usepackage[nolist]{acronym} % Acronym package -- give argument <dua> to suppress acronym usage

%Theorem declarations
\newtheorem{thm}{Theorem}
\newtheorem{lma}{Lemma}
\newtheorem{clm}{Claim}

\newtheoremstyle{dotless}{}{}{\itshape}{}{\bfseries}{}{ }{}
\theoremstyle{dotless}
\newtheorem{defn}{Definition}

\newcommand{\bdefn}[1]{
\begin{defn} {\boldmath (#1)}\textbf{:}
}

\newcommand{\edefn}[1]{
\end{defn}
}

%\icmltitlerunning{Near Optimal Behavior via Approximate State Abstraction}

% --- Notation Commands ---
\newcommand{\ep}{\widetilde \phi}
\newcommand{\epQ}{\ep_{Q^*}}
\newcommand{\epB}{\ep_{\text{bolt}}}
\newcommand{\epMu}{\ep_{\text{mult}}}
\newcommand{\epM}{\ep_{\text{model}}}
\DeclareMathOperator*{\argmin}{arg\,min}
\DeclareMathOperator*{\argmax}{arg\,max}
\newcommand{\mc}[1]{\mathcal{#1}}
\newcommand{\mcS}{\mathcal{S}}
\newcommand{\mcT}{\mathcal{T}}
\newcommand{\mcR}{\mathcal{R}}
\newcommand{\mcA}{\mathcal{A}}
\newcommand{\pM}{\phi_{\text{model}}}

% --- Note Commands ---
\usepackage{color}
\newcommand\dnote[1]{\textcolor{blue}{Dave: #1}}
\newcommand\enote[1]{\textcolor{green}{Ellis: #1}}

%\newcommand\dnote[1]{\textcolor{blue}{\vspace{-1mm}}}
%\newcommand\enote[1]{\textcolor{green}{\vspace{-1mm}}}

% --- DOCUMENT ---
\begin{document}

\twocolumn[
% Suboptimality Bounds for Approximate State Abstraction
\icmltitle{Near Optimal Behavior via Approximate State Abstraction}

\begin{center}
\large{\it To appear at ICML 2016}
\end{center}

% Paper Meta Info
\icmlauthor{David Abel\dagger}{david\_abel@brown.edu}
%\icmladdress{Brown University, 115 Waterman Street, Providence, RI 02906}
\icmlauthor{D. Ellis Hershkowitz\dagger}{david\_hershkowitz@brown.edu}
%\icmladdress{Brown University, 115 Waterman Street, Providence, RI 02906}
\icmlauthor{Michael L.\ Littman}{michael\_littman@brown.edu}
%\icmladdress{Brown University, 115 Waterman Street, Providence, RI 02906}
\icmlkeywords{Reinforcement Learning, State Aggregation, State Abstraction, Planning, MDP}
\vspace{0.3in}
]

%Acronyms -- use \ac for acronym or \acp for plural/capitalization of acronym
\begin{acronym}
\acro{MDP}{Markov Decision Process}
\acrodefplural{MDP}[MDPs]{Markov Decision Processes}
\acro{RL}{reinforcement learning}
\acrodefplural{RL}{Reinforcement learning}
\end{acronym}

% ------------------------
% --- ABSTRACT ---
% ------------------------
\begin{abstract}
\input{abstract.tex}
\end{abstract}

% ---------------------------------------
% --- SECTION: Introduction ---
% ---------------------------------------
\input{introduction.tex}

% ---------------------------------------
% --- SECTION: Background ---
% ---------------------------------------
\input{mdp_background.tex}

% -----------------------------------------
% --- SECTION: Related Work ---
% -----------------------------------------
\input{related_work.tex}

% -----------------------------------------------
% --- SECTION: Abstraction Notion ---
% -----------------------------------------------
\input{abstraction_notation.tex}

% ----------------------------------------------------------------
% --- SECTION: Approximate State Abstraction  ---
% ----------------------------------------------------------------
\input{theory_results.tex}

% -----------------------------------------------
% --- SECTION: Example Domains ---
% -----------------------------------------------
\input{example_domains.tex}

% ------------------------------------------------------------------------------------
% --- SECTION: Approximate Abstraction on Example Domains ---s
% ------------------------------------------------------------------------------------
\input{results_discussion.tex}

% --------------------------------------
% --- SECTION: Conclusion ---
% --------------------------------------
\input{conclusion.tex}

\newpage
% --- BIBLIOGRAPHY ---
\bibliographystyle{icml2016/icml2016}
\bibliography{state_abs}

\end{document}